figures?

### 2023-11-24 Updates/2023-11-17 figures_draft.Rmd
plot bias and sd

Top 10 models and all models coloured by type and showing if they include covariates or not. 
```{r model-type, fig.width=10}
top10_temp <- 
  temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |>
  slice_max(mean_crps, n= 10, na_rm = T) |>
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps)) |> 
  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   colour = "black", 
                   pattern_spacing = 0.02,
                   pattern_alpha = 0.5) + 
  scale_y_reordered() +
  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw()

all_temperature <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps)) |> 
  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   colour = "black", 
                   pattern_spacing = 0.02,
                   pattern_alpha = 0.8) + 
  scale_y_reordered() +
  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw()

ggpubr::ggarrange(top10_temp, all_temperature, nrow = 1, 
                  common.legend = T,
                  legend = 'top', labels = 'auto')

@@

9 of 10 top 10 models use covariates, and process models and ML are the most common best performers. 
There are a lot of really bad empirical models.
## Figure 2
How does model type performance, on avergae, differ by site.

@@

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = model_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'A',) +
  theme_bw()

ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2)
```
Models beeating climatology at the shorter horizons (as expected) PRLA and PRPO always better than climatology.

@@

>>>>

Top 10 models and all models coloured by type and showing if they include covariates or not. 
```{r model-type, fig.width=10}
top10_temp <- 
  temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |>
  slice_max(mean_crps, n= 10, na_rm = T) |>
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   colour = "black", 
                   pattern_spacing = 0.04,
                   pattern_alpha = 1) + 
  scale_y_reordered() +
  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw() + 
    guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
                                  title = 'Uses weather covariates?'),
           fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

all_temp <- temperature_skill |> 
    group_by(model_id) |> 
    summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
    left_join(model_meta) |> 
    mutate(model_id = fct_reorder(model_id, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
    ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
    geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                     stat = "identity",
                     alpha = 0.8, 
                     pattern_fill = "black",
                     colour = "black", 
                     pattern_spacing = 0.03,
                     pattern_alpha = 1) + 
    scale_y_reordered() +
    labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
    geom_vline(xintercept = 0) +
    scale_fill_manual(values = cols_modeltype) +
    scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
    scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
    theme_bw() + 
    guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
           fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, 
                  common.legend = T,
                  legend = 'top', labels = 'auto')

@@

9 of 10 top 10 models use covariates, and process models and ML are the most common best performers. 
There are a lot of really bad empirical models.

## Figure 2
How does model type performance, on avergae, differ by site.

@@

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = model_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw()

ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2)
```

Models beeating climatology at the shorter horizons (as expected) PRLA and PRPO always better than climatology.

@@

### How much is uncertainty?

How does the sd vary - is this my some see a increase in crps skill?

sd <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
  filter(model_id %in% top_mods$model_id, 
         horizon > 0, 
         horizon <=30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_sd = mean(sd, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_sd, colour = model_id)) +
  geom_line()  +  
  labs(y='mean standard deviation', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw()

bias <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime),
         abs_bias = abs(mean - observation)) |> 
  filter(model_id %in% top_mods$model_id, 
         horizon > 0, 
         horizon <=30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_abs_bias, colour = model_id)) +
  geom_line()  +  
  labs(y='mean absolute bias (oC)', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 2.5))

ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2)
```
